---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in FAS_litters data

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


## Take a look at "select()"
Choose some columns and not others.


tell "select()" which column I want to keep.
```{r}
select(litters_df, group, gd0_weight)
```


specify the column range selected
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

to exclude a certain column
```{r}
select(litters_df, -litter_number)
```

renaming columns,
```{r}
select(litters_df, GROUP = group, LitteR_nummber = litter_number)
```
or
```{r}
rename(litters_df, GROUP = group, LittEr_number = litter_number)
```




Select helpers

Select columns that start with "gd"
```{r}
select(litters_df, starts_with("gd"))
```

Select one column and keep everything else.
```{r}
select(litters_df, litter_number, everything())
```
this function does the same thing as the one above (relocating the "litter_number"column to the first)
```{r}
relocate(litters_df, litter_number)
```

## "filter"

```{r}
filter(litters_df, gd0_weight < 22)
```


```{r}
filter(litters_df, gd0_weight >= 22)
```


test for equality, not to assign value to a variable.
```{r}
filter(litters_df, gd_of_birth == 20)
```


"!=" selects values that are not equal to a certain value
```{r}
filter(litters_df, gd_of_birth != 20)
```


select columns that satisfy more than one in a single command
```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```


select row
```{r}
filter(litters_df, group == "Mod8")
```

select collection of rows
```{r}
filter(litters_df, group %in% c("Con7", "Mod8"))
```



## "mutate"
modify/ creat variable

```{r}
litters_df$gd0_weight[litters_df$gd0_weight %in% c("NA", ".", "")] <- NA
litters_df$gd18_weight[litters_df$gd18_weight %in% c("NA", ".", "")] <- NA

litters_df$gd0_weight <- as.numeric(litters_df$gd0_weight)
litters_df$gd18_weight <- as.numeric(litters_df$gd18_weight)

mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight, 
       group = str_to_lower(group))
```

## Arrange

```{r}
arrange(litters_df, pups_born_alive)
```

arrange two columns simultaneously (with order: column "pups_born_alive"first and then "gd0_weight")
```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```
