---
title: "Data Manipulation"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Load in FAS_litters data

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```


## Take a look at "select()"
Choose some columns and not others.


tell "select()" which column I want to keep.
```{r}
select(litters_df, group, gd0_weight)
```


specify the column range selected
```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

to exclude a certain column
```{r}
select(litters_df, -litter_number)
```

renaming columns,
```{r}
select(litters_df, GROUP = group, LitteR_nummber = litter_number)
```
or
```{r}
rename(litters_df, GROUP = group, LittEr_number = litter_number)
```




Select helpers

Select columns that start with "gd"
```{r}
select(litters_df, starts_with("gd"))
```

Select one column and keep everything else.
```{r}
select(litters_df, litter_number, everything())
```
this function does the same thing as the one above (relocating the "litter_number"column to the first)
```{r}
relocate(litters_df, litter_number)
```
