---
title: "Data Import"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
library(haven)
```


## Read in some data

Read in the litters data set.

```{r}
litters_df = read.csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## Take a look at the data

Pringting in the console
```{r}
litters_df
```

```{r}
tail(litters_df)
```
```{r}
skimr::skim(litters_df)
```

## Options to raed_csv

```{r}
litters_df = read_csv("./data/FAS_litters.csv", na = c("", "NA", ".", 999))
# read_csv know there supposed to be missing values, and stored as "."
```

```{r}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
#skip the first 10 rows and rename the column name characters by default
```


check out tread_csv for more info

## otehr file formats

Read an excel file

```{r}
mlb_df = read_excel("./data/mlb11.xlsx")
mlb_df
```

show how the "range =" works
```{r}
mlb_df = read_excel("./data/mlb11.xlsx" , range = "A1:F7")
mlb_df
```

## Read in an SAS file

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with base R
what abt "read.csv"?

```{r}
litters_base = read.csv("./data/FAS_litters.csv")
litters_readr = read_csv("./data/FAS_litters.csv")

litters_base
litters_readr
```

By default "read.csv" print out entire data set, giving more information and the space in column names is change into ".", more difficult to take a look at.